Dismal Code: Studying the Evolution of Security Bugs

### 2013-06-26 docs/evol.tex
final LASER version

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12,TJBD11,M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96,LRWPT97}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse the
relationship of different types of security vulnerabilities to 
evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,HW08,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on security bugs rather than other types of
software defects. This is because compared to other bug categories,
failures due to security bugs have two distinct features:
they can severely affect an organization's infrastructure~\cite{SZ12}, and
they can cause significant financial damage to
an organization~\cite{TH04,BCL08,R06}. 
Specifically,
whereas a software bug can cause a software artifact to fail,
a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain.
In this case, such bugs could give rise to a wide
range of security and privacy issues, like
the access of sensitive information,
the destruction or modification of data, and
denial of service.	
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be to mitigate such bugs.

% @dimitro: say why we did this work, not how we did it
The motivation behind our work was to validate whether programmers care for
the risk posed by security bugs when they release a new version of their software.
In addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug;
in essence, to see whether critical bugs stay unresolved for a long time.
Also, we wanted to elaborate more on the relation of security
bugs with other bug categories.
In the same manner, we tried to examine the relationship
between the size of a project release and the number of security bugs that it contains,
knowing the that research has produced contradictory results on this
issue~\cite{BP84,SYTP85,NBZ06,GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency of
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

@@

\begin{table*}[t]
\centering
\caption{Number of Project Releases that Contain at Least One Security
  High Bug}
\label{tbl:sev}
\leavevmode
	\begin{tabular}{p{31em} r}
	\hline
	Bug Description & Number of Project Releases\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 151\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,579\\
	{\sc pt}: absolute path traversal in servlet  & 103\\
	{\sc pt}: relative path traversal in servlet & 57\\
	{\sc sql}: non-constant string passed to execute method on an {\sc sql} statement & 1,875\\
	{\sc sql}: a prepared statement is generated from a non-constant String & 1,486\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 18\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 90\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 142\\
	\hline
	\end{tabular}
\end{table*}

@@

Another observation involves bugs that we could call {\it
Security High} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection~\cite{RL12} and
Cross-Site Scripting~\cite{WS08}.
To exploit such vulnerabilities, a malicious user does
not have to know anything about the application internals. For all the other
bugs, another program should be written to incorporate references to
mutable objects, access non-final fields, etc.
Also, as bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs coming from the {\it Security High} category,
it might have more vulnerabilities that FindBugs doesn't report.
Table~\ref{tbl:sev} presents the number
of releases where at least one of these bugs exists. In essence, 5,501 releases
($\approx 4,77\% $), contained at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.
The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category
in another subcategory that we call {\it Security Low}.
This categorization was done specifically to point out the
behaviour of bugs that currently top the corresponding lists of most
security providers.\footnote{\url{http://cwe.mitre.org/top25/index.html}}

@@

\subsubsection{The Relation of Defects with the size of a {\sc jar}}

We explored the relation between defects with the size of a project
version, measured by the size of its {\sc jar} file by carrying out
correlation tests between the size and the defect counts for each
project and version. The results, all statistically significant ($p
\ll 0.05$) can be seen in Table~\ref{tbl:jarsizecorr}.  \textbf{The {\it
  Security High} category stands out by having a remarkably lower
effect than the other categories}, even {\it Security Low} that nearly
tops the list.
% dds proposed the following:
Given that, we could say that it would be productive to search for and fix
security bugs even if a project grows bigger.

>>>>

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs, little attention has been paid to study them in relation to
software evolution~\cite{LRWPT97}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse the
relationship of different types of security vulnerabilities to 
evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on security bugs rather than other types of
software defects. This is because compared to other bug categories,
failures due to security bugs have two distinct features:
they can severely affect an organization's infrastructure~\cite{SZ12}, and
they can cause significant financial damage to
an organization~\cite{TH04,BCL08}. 
Specifically,
whereas a software bug can cause a software artifact to fail,
a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain.
In this case, such bugs could give rise to a wide
range of security and privacy issues, like
the access of sensitive information,
the destruction or modification of data, and
denial of service.	
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be to mitigate such bugs.

% @dimitro: say why we did this work, not how we did it
The motivation behind our work was to validate whether programmers care for
the risk posed by security bugs when they release a new version of their software.
In addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug;
in essence, to see whether critical bugs stay unresolved for a long time.
Also, we wanted to elaborate more on the relation of security
bugs with other bug categories.
In the same manner, we tried to examine the relationship
between the size of a project release and the number of security bugs that it contains,
knowing the that research has produced contradictory results on this
issue~\cite{SYTP85,NBZ06,GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency of
a large number of others contains a low rate of security bugs.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

@@

\begin{table*}[t]
\centering
\caption{Number of Project Releases that Contain at Least One ``Security
  High" Bug}
\label{tbl:sev}
\leavevmode
	\begin{tabular}{p{31em} r}
	\hline
	Bug Description & Number of Project Releases\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 151\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,579\\
	{\sc pt}: absolute path traversal in servlet  & 103\\
	{\sc pt}: relative path traversal in servlet & 57\\
	{\sc sql}: non-constant string passed to execute method on an {\sc sql} statement & 1,875\\
	{\sc sql}: a prepared statement is generated from a non-constant String & 1,486\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 18\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 90\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 142\\
	\hline
	\end{tabular}
\end{table*}

@@

Another observation involves bugs that we could call {\it
Security High} and they are a subset of the {\it Security} category.
Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting~\cite{RL12}.
To exploit such vulnerabilities, a malicious user does
not have to know anything about the application internals. For all the other
bugs, another program should be written to incorporate references to
mutable objects, access non-final fields, etc.
Also, as bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs coming from the {\it Security High} category,
it might have more vulnerabilities that FindBugs doesn't report.
Table~\ref{tbl:sev} presents the number
of releases where at least one of these bugs exists. In essence, 5,501 releases
($\approx 4,77\% $), contained at
least one severe security bug. Given the fact that other projects include these
versions as their dependencies, they are automatically rendered vulnerable if
they use the code fragments that include the defects.
The remaining bugs of the {\it Security} category
are grouped together with the bugs of the {\it Malicious Code} category
in another subcategory that we call {\it Security Low}.
This categorization was done specifically to point out the
behaviour of bugs that currently top the corresponding lists of most
security providers.\footnote{\url{http://cwe.mitre.org/top25/index.html}}

@@

\subsubsection{The Relation of Defects with the size of a JAR}

We explored the relation between defects with the size of a project
version, measured by the size of its {\sc jar} file by carrying out
correlation tests between the size and the defect counts for each
project and version. The results, all statistically significant ($p
\ll 0.05$) can be seen in Table~\ref{tbl:jarsizecorr}.  \textbf{The {\it
  Security High} category stands out by having a remarkably lower
effect than the other categories}, even {\it Security Low} that nearly
tops the list.
% dds proposed the following:
%Given that, we could say that it would be productive to search for and fix
%security bugs even if a project grows bigger.
